### 2019-11-23 Mathematical Statatistics/ms_ch_02.rmd
MS ch2, CF ch2
MS ch2, CF ch2
mathematical statistics, ch2
mathematical statistics, ch2

Let the random variable X have a Cauchy distribution with $pdf f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.

@@

>>>>

Let the random variable X have a Cauchy distribution with pdf $f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.

@@

### 2.12

Find:

a.) The $30^{th}$ and $60^{th}$ percentiles for $N(10, 17^2)$.

```{r, echo = T}

# 30th percentile
qnorm(.3, mean = 10, sd = 17)

# 60th percentile
qnorm(.6, mean = 10, sd = 17)

b.) The 0.10 and 0.90 quantile for $N(25, 32^2)$.

```{r, echo = T}

# .1 quantile
qnorm(.1, mean = 25, sd = 32)

# .9 quantile
qnorm(.9, mean = 25, sd = 32)

c.) The point that marks off the upper 25% in $N(25, 32^2)$.

```{r, echo = T}

qnorm(.75, mean = 25, sd = 32, lower.tail = T)

### 2.13

The cdf of the exponential distribution is $F(t) = 1 - e^{-\lambda t}$

a.) Find an expression for the 0.05 quantile $q_{0.05}$.

$q_t = \frac{ln(1 - t)}{-\lambda}$

$q_{0.05} = \frac{0.513}{\lambda}$

b.) Let $\lambda = 4$, and use your answer from (a) to find $q_0.05$, and then check your answer in R using __qexp__.

```{r, echo = T}
qexp(0.05, 4)

-1/4*log(1 - 0.05)

### 2.14

Let $X$ be a random variable with cdf $F(x) = \frac{x^2}{a^2}$ for $0 \le x \le a$.

Find an expression for the $\frac{a}{2}$ and $(1 - \frac{a}{2})$ quantiles, where 0 < a < 1.

$\sqrt{\frac{a}{2}}x$, $1 - \sqrt{\frac{a}{2}}x$

### 2.15

Let X be a random variable with cdf $F(x) = 1 - \frac{9}{x^2}$ for $x \ge 3$.

Find an expression for the $q^{th}$ quantile of $X$.

$\frac{3}{\sqrt{1-x}}$

### 2.16

Let $X \sim {\sf Binom}(20, 0.3)$ and let __F__ denotes its cdf.

Does there exist a __q__ such that $F(q) = 0.05$?

### 2.17

In this exercise, we investigate normal quantile plots using R.

a.) Draw a random sample of size $n = 15$ from $N(0, 1)$, and plot both a normal quantile plot and a histogram.

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample <- function( n ) {
   dat <- data.table( X = rnorm(n) )[, N := .I]

   p1 <- ggplot(dat, aes(sample = X)) +
      geom_qq() +
      geom_qq_line() +
      labs(title = paste0( "Normal Distribution, N=", n ))

   p2 <- ggplot(dat, aes(X)) +
      geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
      geom_density(col = "darkorange", fill = "darkorange", alpha = .3)

   grid.arrange(p1, p2, nrow = 2)
}

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(15)

+ Do the points on the quantile plot appear to fall on a straight line? Is the histogram symmetric, unimodal and bell shaped?

No, the lines do not fit the QQ plot and the histogram is skewed all over.

+ Do this several times.

b.) Repeat part (a) for samples of size n = 30, n = 60, and n = 100.

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(30)

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(60)

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(100)

c.) What lesson do you draw about using graphs to assess whether or not a data set follows a normal distribution?

The random samples from a normal distirbution do not follow the expectation until the sample size is relatively larger (n>30).

### 2.18

Plot by hand the empirical cumulative distribution function for the set of values:

4, 7, 8, 9, 9, 13, 18, 18, 18, 21

### 2.19

The ecdf for a data set with $n = 20$ values is given in figure 2.18.

a.) How many values are less than or equal to 7?

3

b.) How many times does the value 8 occur?

4

c.) In a histogram of these values, how many values fall in the bin (20, 25]?

6

### 2.20

The data set __ChiMarathonMen__ has a sample of times ffor men between 20 and 39 years of age who completed the Chicago Marathon in 2015.

Graph the ecdf's of the times for men in the 25-29-age division and men in the 35-39-age division.

```{r, echo = T, fig.height=4, fig.width=8}
marathon <- data.table(read.csv(paste0(data.dir, "ChiMarathonMen.csv"),
                 header = T))

p1 <- ggplot(marathon[Division == "25-29"], aes(FinishMin)) +
   stat_ecdf(geom = "point") +
   geom_vline(xintercept = 160, color = "cornflowerblue") +
   labs(title = "Men 25-29 Finish Times", y = "")

p2 <- ggplot(marathon[Division == "35-39"], aes(FinishMin)) +
   stat_ecdf(geom = "point") +
   geom_vline(xintercept = 160, color = "cornflowerblue") +
   labs(title = "Men 35-39 Finish Times", y = "")

grid.arrange(p1, p2, nrow = 2)

+ Approximately what proportion of men in these two divisions finished in 160 min or less?

25-29 is around 70%, and 35-39 is around 40%.
